---
title: "Simple Linear Regression Practice"
author: "Alexandra Boville"
date: "2024-03-18"
output: html_document
---

# Practice Simple Linear Regression 


## Load Libraries
```{r}
library(tidyverse)
library(skimr)
library(ggplot2)
library(broom)
library(infer)
```


## Load in Street et al. 2017 dataset
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Street_et_al_2017.csv"

d <- read_csv(f, col_names = TRUE)

head(d)
```

## Some Quick Exploratory Data Analysis
```{r}
skim(d, where(is.numeric))

summary(d)
```

## Step 2: Plotting Data
Plot brain size (ECV) as a function of social group size (Group_size), longevity (Longevity), juvenile period length (Weaning), and reproductive lifespan (Repro_lifespan)
```{r}
p1 <- ggplot(data = d, aes(x = Longevity, y = ECV), na.rm = TRUE) + geom_point()
p2 <- ggplot(data = d, aes(x = Weaning, y = ECV),  na.rm = TRUE) + geom_point()
p3 <- ggplot(data = d, aes(x = Group_size, y = ECV),  na.rm = TRUE) + geom_point()
p4 <- ggplot(data = d, aes(x = Repro_lifespan, y = ECV), na.rm = TRUE) + geom_point()

p1
p2
p3 
p4
```
## Step 3: Derive by Hand regression coeficcients for ECV as a function of Group Size 
```{r}
s <- d %>% 
  filter(!is.na(Group_size) & !is.na(ECV))

(beta1 <- cov(s$ECV, s$Group_size)/var(s$Group_size))

(beta0 <- mean(s$ECV) - beta1 * mean(s$Group_size))
```

# Step 4: Compare Results from Step 3 with lm()

```{r}
m <- lm(ECV ~ Group_size, data = d) #automatically drops missing data
m
```
The values of the regression coefficienct match! 

## Step 5: Repeat Step above for Each Taxonomic Group

creating new variables for each taxonomic group
```{r}
c <- d %>% 
  filter(Taxonomic_group == "Catarrhini") %>% 
  filter(!is.na(Group_size) & !is.na(ECV))

P <- d %>% 
  filter(Taxonomic_group == "Platyrrhini") %>% 
  filter(!is.na(Group_size) & !is.na(ECV))

S <- d %>% 
  filter(Taxonomic_group == "Strepsirhini") %>% 
  filter(!is.na(Group_size) & !is.na(ECV))
```

Determining regression coefficients 
```{r}
#calculating regression coefficients for Catarrhini
(beta1_c <- cov(c$ECV, c$Group_size)/var(c$Group_size))
(beta0_c <- mean(c$ECV) - beta1 * mean(c$Group_size))

#calculating regression coefficients for Platyrrhini
(beta1_P <- cov(P$ECV, P$Group_size)/var(P$Group_size))
(beta0_P <- mean(P$ECV) - beta1 * mean(P$Group_size))

#calculating regression coefficients for Strepsirhini
(beta1_S <- cov(S$ECV, S$Group_size)/var(S$Group_size))
(beta0_S <- mean(S$ECV) - beta1 * mean(S$Group_size))
```
The regression coefficients differ for each taxonomic group, mostly in the beta0 value (intercept). To fully confirm the degree of difference it would be beneficial to determine the standard error and if these coefficents are statistically significant. 

## Step 6:Calculating the standard error for the slope coefficient, the 95% CI, and the p value associated with this coefficient 
Determining SE
```{r}
# Standard Error of slope coefficient
df_regression <- 1  # p = 1
(df_error <- nrow(s) - df_regression - 1 )# n - p - 1
(df_y <- nrow(s) - df_regression)

#SSE = height - predicted height
SSE <- sum(m$residuals^2)
SSE
#SSR = predicted height - mean height
SSR <- sum((m$fitted.values - mean(m$model$Group_size))^2)
SSR

(MSR <- SSR/df_regression)  # mean variance explained by the regression equation
(MSE <- SSE/df_error)  # mean remaining variance

(SSX <- sum((s$Group_size - mean(s$Group_size))^2)) # how much x variation there is
SEbeta1 <- sqrt(MSE/SSX)
SEbeta1

```
I'm not sure why I am getting NA's for these values... I already removed NA's from this data previously...

95% CI by hand
```{r}
alpha <- 0.05
m_aug <- augment(m, se_fit = TRUE)
head(m_aug)


m_aug <- m_aug %>% 
  mutate(c.lwr = .fitted - qt(1 - alpha/2, nrow(m_aug) - 2) * .se.fit, c.upr = .fitted +
    qt(1 - alpha/2, nrow(m_aug) - 2) * .se.fit)
head(m_aug)

```


p-value by hand
```{r}
# fratio = the ratio of the variance explained by the regression model to the remaining, unexplained variance (MSR/MSE)
fratio <- MSR/MSE
fratio

pf(q = fratio, df1 = 1, df2 = 153, lower.tail = FALSE)
```

Determing SE, 95% CI and p-value by using lm()
```{r}
m_c <- lm(ECV ~ Group_size, data = c) 
m_P <- lm(ECV ~ Group_size, data = P) 
m_S <- lm(ECV ~ Group_size, data = S) 

tidy(m_c)
tidy(m_P)
tidy(m_S)

(CI <- confint(m, level = 1 - alpha))

```


## Step 7: Using a Permutation Approach
Use a permutation approach with 1000 permutations to generate a null sampling distribution for the slope coefficient.

```{r}
#Defining alpha, CI boundaries, and critical values
alpha <- 0.05
confidence_level <- 1 - alpha
con_lower <- alpha/2
con_upper <- 1 - (alpha/2)
degrees_of_freedom <- nrow(d) - 2
critical_value <- qt(con_upper, df = degrees_of_freedom)

# original slope
og.slope <- lm(data = d, ECV ~ Group_size) %>%
    # tidy the model and add the CI based on the t distribution
  tidy(conf.int = TRUE, conf.level = confidence_level) %>%
    # or manually calculate the CI based on the t distribution
  mutate(lower = estimate - std.error * critical_value, upper = estimate + std.error *
    critical_value) %>%
    filter(term == "Group_size")
og.slope 

permuted.slope <- d %>%
  specify(ECV ~ Group_size) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")

head(permuted.slope) 

get_ci(permuted.slope, level = 1 - alpha, type = "percentile")

p.value <- permuted.slope %>% 
  mutate(abs_stat=abs(stat)) %>%
  summarize(estimate = mean(abs_stat >= abs(pull(og.slope, estimate)))
  )
p.value
```
## Step 8: Using Bootstrapping to Generate a CI 
Use bootstrapping to generate a 95% CI for your estimate of the slope coefficient using both the percentile method and the theory-based method

Creating Confidence Intervals
```{r}
boot.slope <- d %>%
  specify(ECV ~ Group_size) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "slope")

head(boot.slope) 
```

Percentile Method
```{r}
CI.percentile <- get_ci(boot.slope, level = 1 - alpha, type = "percentile")
CI.percentile
```

Theory-based Method
```{r}

boot.slope.summary <- boot.slope %>%
  summarize(
    estimate = mean(stat),
    std.error = sd(stat),
    boot.lower = quantile(stat, con_lower),
    boot.upper = quantile(stat, con_upper)
  )

boot.slope.summary


CI.theory <- get_ci(boot.slope, level = 1 - alpha, type = "se", point_estimate = pull(boot.slope.summary,
    estimate))
CI.theory
```



